Orientation Control System: Enhancing Aerial Maneuvers for Quadruped Robots

### 2023-01-05 answersToReviewers.tex

%  LaTeX support: latex@mdpi.com 
%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.

%=================================================================
\documentclass[sensors,reply,submit,pdftex,moreauthors]{Definitions/mdpi} 
% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept". The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, adolescents, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, allergies, alloys, analytica, animals, antibiotics, antibodies, antioxidants, applbiosci, appliedchem, appliedmath, applmech, applmicrobiol, applnano, applsci, aquacj, architecture, arts, asc, asi, astronomy, atmosphere, atoms, audiolres, automation, axioms, bacteria, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomass, biomechanics, biomed, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, blsf, brainsci, breath, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, chips, cimb, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coasts, coatings, colloids, colorants, commodities, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, covid, crops, cryptography, crystals, csmf, ctn, curroncol, currophthalmol, cyber, dairy, data, dentistry, dermato, dermatopathology, designs, diabetology, diagnostics, dietetics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, ent, entomology, entropy, environments, environsciproc, epidemiologia, epigenomes, est, fermentation, fibers, fintech, fire, fishes, fluids, foods, forecasting, forensicsci, forests, foundations, fractalfract, fuels, futureinternet, futureparasites, futurepharmacol, futurephys, futuretransp, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, hazardousmatters, healthcare, hearts, hemato, heritage, highthroughput, histories, horticulturae, humanities, humans, hydrobiology, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijtm, ijtpp, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jal, jcdd, jcm, jcp, jcs, jdb, jeta, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jzbg, kidney, kidneydial, knowledge, land, languages, laws, life, liquids, literature, livers, logics, logistics, lubricants, lymphatics, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, merits, metabolites, metals, meteorology, methane, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, microplastics, minerals, mining, modelling, molbank, molecules, mps, msf, mti, muscles, nanoenergyadv, nanomanufacturing, nanomaterials, ncrna, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, nri, nursrep, nutraceuticals, nutrients, obesities, oceans, ohbm, onco, oncopathology, optics, oral, organics, organoids, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacoepidemiology, pharmacy, philosophies, photochem, photonics, phycology, physchem, physics, physiologia, plants, plasma, pollutants, polymers, polysaccharides, poultry, powders, preprints, proceedings, processes, prosthesis, proteomes, psf, psych, psychiatryint, psychoactives, publications, quantumrep, quaternary, qubs, radiation, reactions, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, rheumato, risks, robotics, ruminants, safety, sci, scipharm, seeds, sensors, separations, sexes, signals, sinusitis, skins, smartcities, sna, societies, socsci, software, soilsystems, solar, solids, sports, standards, stats, stresses, surfaces, surgeries, suschem, sustainability, symmetry, synbio, systems, taxonomy, technologies, telecom, test, textiles, thalassrep, thermo, tomography, tourismhosp, toxics, toxins, transplantology, transportation, traumacare, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, venereology, vetsci, vibration, viruses, vision, waste, water, wem, wevj, wind, women, world, youth, zoonoticdis

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by: 
% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
% MDPI internal commands
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2022}
\copyrightyear{2022}
%\externaleditor{Academic Editor: Firstname Lastname}
\datereceived{} 
%\daterevised{} % Only for the journal Acoustics
\dateaccepted{} 
\datepublished{} 
%\datecorrected{} % Corrected papers include a "Corrected: XXX" date in the original paper.
%\dateretracted{} % Corrected papers include a "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/} % If needed use \linebreak
%\doinum{}
%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption
\usepackage{glossaries}
\usepackage{bm}
\usepackage{tensor}
%\input{includes/used_packages.tex}
\input{includes/acronyms.tex}
%\input{includes/math_commands.tex}
%\input{includes/newcommands.tex}

\newcommand{\MF}[1]{\textcolor{red}{\textbf{mfocchi}: #1}}
\newcommand{\CS}[1]{\textcolor{violet}{\textbf{csemini}: #1}}
\newcommand{\FR}[1]{\textcolor{teal}{\textbf{froscia}: #1}}
\newcommand{\ADP}[1]{\textcolor{blue}{\textbf{adelprete}: #1}}
%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Orientation Control System: Enhancing Aerial Maneuvers for Quadruped Robots}

% MDPI internal command: Title for citation in the left column
\TitleCitation{Orientation Control System: Enhancing Aerial Maneuvers for Quadruped Robots}

% Author Orchid ID: enter ID or remove command
%\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Francesco Roscia $^{1}$, Andrea Cumerlotti $^{1, \, 2}$, Andrea Del Prete $^{2}$, Claudio Semini $^{1}$ and Michele Focchi$^{1, \, 3, ^*}$}

%\longauthorlist{yes}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Francesco Roscia, Andrea Cumerlotti, Andrea Del Prete, Claudio Semini and Michele Focchi}

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Roscia, F.; Cumerlotti, A.; Del Prete, A., Semini, C.; Focchi, M.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Dynamic Legged Systems (DLS) lab, Istituto Italiano di Tecnologia (IIT), Genoa, Italy\\
$^{2}$ \quad Industrial Engineering Department (DII), University of Trento, Trento, Italy\\
$^{3}$ \quad Department of Information Engineering and Computer Science (DISI), University of Trento, Trento, Italy\\
}

% Contact information of the corresponding author
\corres{Correspondence: michele.focchi@unitn.it}

% Current address and/or shared authorship
%\firstnote{Current address: Affiliation 3.} 
%\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

%\simplesumm{} % Simple summary

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\) 
\makeglossaries
% Keywords

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data
%\dataset{DOI number or link to the deposited data set if the data set is published separately. If the data set shall be published as a supplement to this paper, this field will be filled by the journal editors. In this case, please submit the data set as a supplement.}
%\datasetlicense{License under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Encyclopedia
%\encyclopediadef{For entry manuscripts only: please provide a brief overview of the entry title instead of an abstract.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Advances in Respiratory Medicine
%\addhighlights{yes}
%\renewcommand{\addhighlights}{%

%\noindent This is an obligatory section in “Advances in Respiratory Medicine”, whose goal is to increase the discoverability and readability of the article via search engines and other scholars. Highlights should not be a copy of the abstract, but a simple text allowing the reader to quickly and simplified find out what the article is about and what can be cited from it. Each of these parts should be devoted up to 2~bullet points.\vspace{3pt}\\
%\textbf{What are the main findings?}
% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
% \item First bullet.
% \item Second bullet.
% \end{itemize}\vspace{3pt}
%\textbf{What is the implication of the main finding?}
% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
% \item First bullet.
% \item Second bullet.
% \end{itemize}
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Dear Ms. Vera Popovic,\\
Thank you for giving us the chance to submit a revised draft version of the manuscript titled "Orientation Control System: Enhancing Aerial Maneuvers for Quadruped Robots" to the MDPI journal \textit{Sensors}. We are grateful for the time and the effort you and the reviewers have dedicated to providing insightful comments on the original manuscript. We have incorporated changes to reflect most of the provided suggestions: they are highlighted in red within the manuscript. For removing the highlights, please comment line 73 and uncomment line 74 in the file named \texttt{mdpi2022roscia.tex}.\\
Here is a point-by-point response to the reviewers' comments and concerns.

\section*{Comments for Reviewer 1}
\begin{itemize}
	\item \textbf{Comment 1:} \textit{Extend the text of the manuscript (e.g. introduction or conclusion) with specific results in the world and Europe, - Improve the quality of the paper by presenting the results of publications by researchers and experts who are involved in this field and registered in world databases (wos). These are e.g: Measurement of industrial robot pose repeatability, Investigation of snake robot locomotion possibilities in a pipe, thanks.}
	\item[] \textbf{Response:} Thank you for this advice. However, in the case of our study, it seems slightly out of range because we tackle of the angular dynamics of the main body of a quadruped robot, that is observable.	\\
	\item \textbf{Comment 2:} \textit{Figure 4 should be contrasting and readable}
	\item[] \textbf{Response:} Thank you for pointing this out. We have incorporated your suggestion in the revised manuscript. \\
	\item \textbf{Comment 3:} \textit{ Conclusions and future work should be extended to contain practical applications based on research described in this paper - expand references.}
	\item[] \textbf{Response:} We are glad you made such an important hint to improve the quality of the manuscript. Possible applications of the presented OCS include but are not limited to efficiently adjust the posture of quadruped robots walking or jumping on uneven terrains. As proved in the third simulation (back-flip), our approach improves the capabilities of quadrupeds in space application enabling fast locomotion by means of leaps, ensuring a reactive control action on the robot angular momentum. Furthermore, the method presented in Section 3 for designing the OCS does not depend on the specific platform, thus it can be replicated for reorienting mechanical structures with different morphology, e.g, monopods or bipeds. A construction worker's backpack can contain two flywheels with incident rotation axes: in the event of a fall from scaffolding, they can be used to reorient the human body to avoid impact of the head with the ground with the same controller proposed in this work. \\
	\item \textbf{Comment 4:} \textit{Number all mathematical equations}
	\item[] \textbf{Response:} Thank you for bringing this to our attention. We applied the comment into the document.\\	
	\item \textbf{Comment 5:} \textit{For article type, 20 references are not enough. Please add more references (>20) during your revisions.}
	\item[] \textbf{Response:} We really appreciate your suggestion, which helped us to improve the manuscript. We included other relevant works in the literature review, e.g. [An, 2022], [Tang, 2022], and [Kurtz, 2022].\\
	\item \textbf{Comment 6:} \textit{Standardization of literature input (into one style)}
	\item[] \textbf{Response:} Agree. We have corrected the literature input.\\
\end{itemize}

### 2023-01-05 mdpi2022roscia.tex
updated pdf
update after reviews

\newcommand{\MF}[1]{\textcolor{red}{\textbf{mfocchi}: #1}}
\newcommand{\CS}[1]{\textcolor{violet}{\textbf{csemini}: #1}}
\newcommand{\FR}[1]{\textcolor{teal}{\textbf{froscia}: #1}}
\newcommand{\ADP}[1]{\textcolor{blue}{\textbf{adelprete}: #1}}
%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

@@

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{For legged robots, aerial motions are the only option to overpass obstacles that cannot be circumvent with standard locomotion gaits. In these cases, the robot must perform a leap to either jump onto the obstacle or fly over it. However, these movements represent a challenge because during the flight phase the \gls{com} cannot be controlled, and the robot orientation has limited controllability. This paper focuses on the latter issue and proposes an \gls{ocs} consisting of two rotating and actuated masses (flywheels or reaction wheels) to gain control authority on the robot orientation. Because of the conservation of angular momentum, their rotational velocity can be adjusted to steer the robot orientation even when there are no contacts with the ground. The axes of rotation of the flywheels are designed to be incident, leading to a compact orientation control system that is capable of controlling both roll and pitch angles, considering the different moment of inertia in the two directions. We tested the concept with simulations on the robot Solo12.}

@@

\begin{figure}[t!]
	\centering
	\includegraphics[width=7cm,height=8cm,keepaspectratio]{figures/solo_fw_pic_compressed.png}
	\caption{\small The proposed Orientation Control System for 
		for the $2.5 \ \mathrm{kg}$ quadruped robot Solo12 consists of two $0.1 \ \mathrm{kg}$ flywheels with incident rotation axes. Each wheel is located in a 3D-printed shell and mounted on the trunk body.}
	\label{fig:solo12flywheels}
\end{figure}

@@

\subsection{Proposed Approach and Contribution}
In this work, we \MF{introduce} a compact \gls{ocs} based on two flywheels, mounted on the trunk of the lightweight quadruped Solo12 \cite{grimminger2020open}. \MF{The open-source hardware and software are reported in \cite{github_solo}. Fig. \ref{fig:solo12flywheels} shows the \gls{ocs} mounted on the trunk of the robot. }
The contributions of the paper are:
\begin{itemize}
	\item the design of a novel \gls{ocs} that enables to control the orientation of a legged robot during a jump in an effective way, while keeping the design simple. 
	In particular, the axes of rotation of the flywheels are set to be incident, enabling \textit{continuous} controllability in both the directions (roll and pitch) while keeping the device compact.
	\item simulations  with the quadruped Solo12 that demonstrate the effectiveness of the proposed approach.
\end{itemize}

@@

\section{Orientation Control System}
\label{sec:ocs}
In this section, we present a procedure to select the inertia of the flywheels, and we show how to exploit them for simultaneously controlling the robot roll and pitch orientation.

\subsection{Bounds on the Inertia}
Investigation for the selection of the flywheels' inertia can be performed with the Elroy's Beanie model, depicted in Fig. \ref{fig:elroy-beanie}.
\begin{figure}
	\centering
	\includegraphics[width=.7\linewidth]{figures/elroys-beanie.eps}
	\caption{\small Schematic representation of the Elroy's Beanie model used for the preliminary analysis of the pitch motion.}
	\label{fig:elroy-beanie}
\end{figure} 
\begin{figure}
	\centering
	\includegraphics[width=.8\linewidth]{figures/lowerbound2.eps}
	\caption{\small Minimum inertia of the \gls{ocs} (normalized by the robot inertia) necessary to achieve a base velocity variation $\Delta \dot{\theta}$, given the actuation bound $\dot \gamma^{max}$.}
	\label{fig:inertia_limits}
\end{figure}
This consists of two rigid bodies connected through their \gls{com} with a revolute joint. One of the bodies represents the robot in its nominal configuration. The other models the \gls{ocs}, here represented by a single wheel for the sake of simplicity. The aim is to examine the rotational dynamics of the system as a whole. To simplify the analysis, in the following we will refer only to the effects on the pitch angle of the robot, keeping in mind that the same arguments apply also to the roll. 
Let us identify the moment of inertia of the robot in the nominal configuration as $I_r$ and the moment of inertia of the two flywheels as $I_f$, both referred to the system \gls{com}.
The angular momentum $L$ of this system can be written as
\begin{equation}\label{eq:ang_mom_elroy}
L(t) = \left(I_r + I_{f}\right)\dot{\theta}(t) + I_{f}\dot{\gamma}(t)
\end{equation}
where $\dot{\theta}$ and $\dot{\gamma}$ are respectively the robot pitch rate and the angular speed of the wheel.
We want to drive the robot to a desired pitch rate $\dot{\theta}_{des}$ by acting on the wheel speed.
Without loss of generality, \MF{one} can assume the flywheel is stationary with the robot at the instant at which the reorientation maneuver starts ($\dot{\gamma}_0 = 0$), having a system angular momentum of $L_0 = \left(I_r + I_{f}\right)\dot{\theta}_0$.
Under the condition of conservation of the angular momentum, this quantity is constant over time and it is possible to estimate the lower bound for $I_{f}$ given a desired pitch rate and the maximum velocity of the flywheels $\dot{\gamma}_{max}$:
\begin{equation}
I_{f} \geq I_r \dfrac{ \Delta \dot{\theta} }{\Delta \dot{\theta}+\dot{\gamma}_{max}}.
\end{equation}
where $\Delta \dot{\theta} =  \left\lvert \dot{\theta}_0-\dot{\theta}_{des} \right\rvert$ is the base velocity variation.
Fig. \ref{fig:inertia_limits} reports the lower bound of $I_{f}$ given the desired base velocity variation and the maximum velocity of the actuator.

@@

To remove unnecessary complications, we designed two identical modules, to be mounted on the trunk of Solo12, symmetrically with the respect of the robot sagittal plane. We denote with $\alpha \leq \pi/2$ the non-negative incident angle between the wheels' axes of rotation and the robot lateral direction. The matrix 

@@

\subsection{Inertia Selection}
Once the desired inertia $I_f$ is chosen, according to the 2D Elroy's Beanie model, it is used to realize the 3D \gls{ocs} (Fig. \ref{fig:flywheel}). 
We decided to design the flywheels in the shape of hollow cylinders. This shape increases the inertia by locating the mass far away from the rotation axis. The inertia tensor expressed in its principal axes, $\bm{I}_f = \mathrm{diag}\left\{ I_{f, xx}, \, I_{f, yy}, \, I_{f, zz}  \right\}$, depends on the cylinders inner and outer radii, $r$ and $R$, its height $h$ and the material density $\rho$:
\begin{equation}
\begin{split}
I_{f, xx} = I_{f, yy} &= \frac{1}{12}\pi \rho h \left(3\left(R^4 - r^4\right) + h^2\left(R^2 - r^2\right)\right) \\
I_{f, zz} &= \frac{1}{2}\pi \rho h \left(R^4 - r^4\right)
\end{split}
\end{equation}
Notice that in the Elroy's Beanie model, there is a single body that models the complete \gls{ocs}. The inertia introduced in the Elroy's Beanie model has to be split between the two flywheels: $I_{f, zz} = I_f/\left(2\cos(\alpha)\right)$. 
The parameters $R$ and $h$ can be set to have a compact \gls{ocs} and $\rho$ depends on the chosen material, that is stainless steel in our case. The inner radius $r$ can be adjusted to obtain the desired inertia. Spokes with negligible mass connect the wheel to the motor shaft. All the parameters are reported in Table \ref{tab:dim_fw}, together with the selected inertia and mass of a single flywheel.
\begin{table}[H] 
	\caption{Sizes and dynamic parameters of a single flywheel}
	\label{tab:dim_fw}
	\newcolumntype{C}{>{\centering\arraybackslash}X}
	\begin{tabularx}{\textwidth}{CCC}
		\toprule
		\textbf{Parameter} & \textbf{Value} & \textbf{Unit}\\
		\midrule
		$r$ & $2.20\cdot 10^{-2}$ & $\mathrm{m}$ \\
		$R$ & $3.00\cdot 10^{-2}$ & $\mathrm{m}$ \\
		$h$   & $1.02\cdot 10^{-2}$ & $\mathrm{m}$ \\
		$\rho$   & $7.86 \cdot 10^{+3}$ & $\mathrm{kg/m^3}$ \\
		$m$   & $1.02\cdot 10^{-1}$ & $\mathrm{kg}$ \\
		$I_{xx}, I_{yy}$   & $3.64 \cdot 10^{-5}$ & $\mathrm{kg \cdot m^2}$ \\
		$I_{zz}$   & $7.11 \cdot 10^{-5}$ & $ \mathrm{kg \cdot m^2}$\\
		\bottomrule	\end{tabularx}
\end{table}
\begin{figure}[t!]
	\centering
	\includegraphics[width=7cm,height=8cm,keepaspectratio]{figures/flywheel_opaque.png}
	\caption{\small The final design of the flywheel, shown together with its inertia principal axes. Here, $h$ is the wheel thickness, $r$ is the inner radius of the wheel and $R$ is the outer one.}
	\label{fig:flywheel}
\end{figure}
\unskip

>>>>

\newcommand{\MF}[1]{\textcolor{red}{#1}}
%\newcommand{\MF}[1]{\textcolor{black}{#1}}
\newcommand{\CS}[1]{\textcolor{violet}{\textbf{csemini}: #1}}
\newcommand{\FR}[1]{\textcolor{teal}{\textbf{froscia}: #1}}
\newcommand{\ADP}[1]{\textcolor{blue}{\textbf{adelprete}: #1}}
%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

@@

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{For legged robots, aerial motions are the only option to overpass obstacles that cannot be circumvent with standard locomotion gaits. In these cases, the robot must perform a leap to either jump onto the obstacle or fly over it. However, these movements represent a challenge because during the flight phase the \gls{com} cannot be controlled, and the robot orientation has limited controllability. This paper focuses on the latter issue and proposes an \gls{ocs} consisting of two rotating and actuated masses (flywheels or reaction wheels) to gain control authority on the robot orientation. Because of the conservation of angular momentum, their rotational velocity can be adjusted to steer the robot orientation even when there are no contacts with the ground. The axes of rotation of the flywheels are designed to be incident, leading to a compact orientation control system that is capable of controlling both roll and pitch angles, considering the different moment of inertia in the two directions. \MF{The concept is tested with simulations on the robot Solo12.}}

@@

@@

\subsection{Proposed Approach and Contribution}
\begin{figure}[t!]
	\centering
	\includegraphics[width=7cm,height=8cm,keepaspectratio]{figures/solo_fw_pic_compressed.png}
	\caption{\small The proposed Orientation Control System for 
		for the $2.5 \ \mathrm{kg}$ quadruped robot Solo12 consists of two $0.1 \ \mathrm{kg}$ flywheels with incident rotation axes. Each wheel is located in a 3D-printed shell and mounted on the trunk body.}
	\label{fig:solo12flywheels}
\end{figure} 
\MF{This work introduces} a compact \gls{ocs} based on two flywheels, mounted on the trunk of the lightweight quadruped Solo12 \cite{grimminger2020open}. \MF{The open-source hardware and software are reported in \cite{github_solo}. Fig. \ref{fig:solo12flywheels} shows the \gls{ocs} mounted on the trunk of the robot. }
The contributions of the paper are:
\begin{itemize}
	\item the design of a novel \gls{ocs} that enables to control the orientation of a legged robot during a jump in an effective way, while keeping the design simple. 
	In particular, the axes of rotation of the flywheels are set to be incident, enabling \textit{continuous} controllability in both the directions (roll and pitch) while keeping the device compact.
	\item simulations  with the quadruped Solo12 that demonstrate the effectiveness of the proposed approach.
\end{itemize}

@@

\section{Orientation Control System}
\label{sec:ocs}
In this section, a procedure to select the inertia of the flywheels \MF{is presented, together with a strategy} to exploit them for simultaneously controlling the robot roll and pitch orientation.

\subsection{Bounds on the Inertia}
Investigation for the selection of the flywheels' inertia can be performed with the Elroy's Beanie model, depicted in Fig. \ref{fig:elroy-beanie}.
\begin{figure}[t!]
	\centering
	\includegraphics[width=.7\linewidth]{figures/elroys-beanie.eps}
	\caption{\small Schematic representation of the Elroy's Beanie model used for the preliminary analysis of the pitch motion.}
	\label{fig:elroy-beanie}
\end{figure} 
\begin{figure}[t]
	\centering
	\includegraphics[width=.8\linewidth]{figures/lowerbound2.eps}
	\caption{\small Minimum inertia of the \gls{ocs} (normalized by the robot inertia) necessary to achieve a base velocity variation $\Delta \dot{\theta}$, given the actuation bound $\dot \gamma^{max}$.}
	\label{fig:inertia_limits}
\end{figure}
This consists of two rigid bodies connected through their \gls{com} with a revolute joint. One of the bodies represents the robot in its nominal configuration. The other models the \gls{ocs}, here represented by a single wheel for the sake of simplicity. The aim is to examine the rotational dynamics of the system as a whole. To simplify the analysis, in the following \MF{only the effects on the pitch angle of the robot is considered}, keeping in mind that the same arguments apply also to the roll. 
Let us identify the moment of inertia of the robot in the nominal configuration as $I_r$ and the moment of inertia of the two flywheels as $I_f$, both referred to the system \gls{com}.
The angular momentum $L$ of this system can be written as
\begin{equation}\label{eq:ang_mom_elroy}
L(t) = \left(I_r + I_{f}\right)\dot{\theta}(t) + I_{f}\dot{\gamma}(t)
\end{equation}
where $\dot{\theta}$ and $\dot{\gamma}$ are respectively the robot pitch rate and the angular speed of the wheel.
\MF{The robot is driven to reach} a desired pitch rate $\dot{\theta}_{des}$ by acting on the wheel speed.
Without loss of generality, \MF{one} can assume the flywheel is stationary with the robot at the instant at which the reorientation maneuver starts ($\dot{\gamma}_0 = 0$), having a system angular momentum of $L_0 = \left(I_r + I_{f}\right)\dot{\theta}_0$.
Under the condition of conservation of the angular momentum, this quantity is constant over time and it is possible to estimate the lower bound for $I_{f}$ given a desired pitch rate and the maximum velocity of the flywheels $\dot{\gamma}_{max}$:
\begin{equation}
I_{f} \geq I_r \dfrac{ \Delta \dot{\theta} }{\Delta \dot{\theta}+\dot{\gamma}_{max}}.
\end{equation}
where $\Delta \dot{\theta} =  \left\lvert \dot{\theta}_0-\dot{\theta}_{des} \right\rvert$ is the base velocity variation.
Fig. \ref{fig:inertia_limits} reports the lower bound of $I_{f}$ given the desired base velocity variation and the maximum velocity of the actuator.

@@

To remove unnecessary complications, \MF{the authors} designed two identical modules, to be mounted on the trunk of Solo12, symmetrically with the respect of the robot sagittal plane. \MF{Let} denote with $\alpha \leq \pi/2$ the non-negative incident angle between the wheels' axes of rotation and the robot lateral direction. The matrix 

@@

\subsection{Inertia Selection}
Once the desired inertia $I_f$ is chosen, according to the 2D Elroy's Beanie model, it is used to realize the 3D \gls{ocs} (Fig. \ref{fig:flywheel}). 
\MF{Two identical flywheels are designed to have} the shape of hollow cylinders. This shape increases the inertia by locating the mass far away from the rotation axis. The inertia tensor expressed in its principal axes, $\bm{I}_f = \mathrm{diag}\left\{ I_{f, xx}, \, I_{f, yy}, \, I_{f, zz}  \right\}$, depends on the cylinders inner and outer radii, $r$ and $R$, its height $h$ and the material density $\rho$:
\begin{equation}
\begin{split}
I_{f, xx} = I_{f, yy} &= \frac{1}{12}\pi \rho h \left(3\left(R^4 - r^4\right) + h^2\left(R^2 - r^2\right)\right) \\
I_{f, zz} &= \frac{1}{2}\pi \rho h \left(R^4 - r^4\right)
\end{split}
\end{equation}
Notice that in the Elroy's Beanie model, there is a single body that models the complete \gls{ocs}. The inertia introduced in the Elroy's Beanie model has to be split between the two flywheels: $I_{f, zz} = I_f/\left(2\cos(\alpha)\right)$. 
The parameters $R$ and $h$ can be set to have a compact \gls{ocs} and $\rho$ depends on the chosen material, that is stainless steel in our case. The inner radius $r$ can be adjusted to obtain the desired inertia. Spokes with negligible mass connect the wheel to the motor shaft. All the parameters are reported in Table \ref{tab:dim_fw}, together with the selected inertia and mass of a single flywheel.
\begin{figure}[t!]
	\centering
	\includegraphics[width=7cm,height=8cm,keepaspectratio]{figures/flywheel_opaque.png}
	\caption{\small The final design of the flywheel, shown together with its inertia principal axes. Here, $h$ is the wheel thickness, $r$ is the inner radius of the wheel and $R$ is the outer one.}
	\label{fig:flywheel}
\end{figure}
\begin{table}[t] 
	\caption{Sizes and dynamic parameters of a single flywheel}
	\label{tab:dim_fw}
	\newcolumntype{C}{>{\centering\arraybackslash}X}
	\begin{tabularx}{\textwidth}{CCC}
		\toprule
		\textbf{Parameter} & \textbf{Value} & \textbf{Unit}\\
		\midrule
		$r$ & $2.20\cdot 10^{-2}$ & $\mathrm{m}$ \\
		$R$ & $3.00\cdot 10^{-2}$ & $\mathrm{m}$ \\
		$h$   & $1.02\cdot 10^{-2}$ & $\mathrm{m}$ \\
		$\rho$   & $7.86 \cdot 10^{+3}$ & $\mathrm{kg/m^3}$ \\
		$m$   & $1.02\cdot 10^{-1}$ & $\mathrm{kg}$ \\
		$I_{xx}, I_{yy}$   & $3.64 \cdot 10^{-5}$ & $\mathrm{kg \cdot m^2}$ \\
		$I_{zz}$   & $7.11 \cdot 10^{-5}$ & $ \mathrm{kg \cdot m^2}$\\
		\bottomrule	\end{tabularx}
\end{table}
\unskip
